### 2017-07-03 isolation.tex
Isolation.

Operational semantics of Fig.~\ref{fig:txnimp} does not employ
concurrency control constructs in any form. Any additional concurrency
control provided by the database implementation, beyond that which is
already incuded in the isolation specification, also needs to be
axiomatized in $\I$. We first describe such axiomatizations, and then
present the specifications of the standard isolation levels as
implemented by popular off-the-shelf databases.

@@

@@

>>>>

Databases implement isolation levels through a combination of various
concurrency control mechanisms, which are absent from
Fig.~\ref{fig:txnimp}. In this section, we first axiomatize the
semantics of various concurrency control constructs, and then define
the precise semantics of the isolation levels as they are implemented
on Postgres and MySQL datbases.
% Any additional concurrency control provided by the database
% implementation, beyond that which is already incuded in the
% isolation specification, also needs to be axiomatized in $\I$. We
% first describe such axiomatizations, and then present the
% specifications of the standard isolation levels as implemented by
% popular off-the-shelf databases.

@@

\textbf{Snapshots} Both MySQL and Postgres implement isolation levels
that execute transactions against a static snapshot of the database.
The semantics of a snapshot be axiomatized as following:
\begin{smathpar}
\begin{array}{lcl}
  \I_{ss}(\stl,\stg,\stg') & = & \stg' = \stg.
\end{array}
\end{smathpar}
% The snapshot semantics however may not be applicable throughout the
% lifetime of the transaction, like in the case of Repeatable Read (RR).

\textbf{Rollbacks} Rollbacks triggered by the post-facto detection of
write-write conflicts can be axiomatized as following:
\begin{smathpar}
\begin{array}{lcl}
  \I_{rb}(\stl,\stg,\stg') & = & \forall(r\in\stl)(r'\in\stg).~
  r'.\idf = r.\idf \Rightarrow r'\in\stg'.
\end{array}
\end{smathpar}

@@

%\usepackage[table,xcdraw]{xcolor}
\begin{table}[]
\centering
\begin{tabular}{l|l|l|l|l|l|l|l|l|l|l|l|l|}
\cline{2-13}
                                & \multicolumn{6}{c|}{{\color[HTML]{333333} Postgres}}                         & \multicolumn{6}{c|}{MySql}                                                   \\ \cline{2-13} 
                                & \multicolumn{2}{c|}{RC} & \multicolumn{2}{c|}{RR} & \multicolumn{2}{c|}{SER} & \multicolumn{2}{c|}{RC} & \multicolumn{2}{c|}{RR} & \multicolumn{2}{c|}{SER} \\ \cline{2-13} 
                                & T          & F          & T          & F          & T           & F          & T          & F          & T          & F          & T           & F          \\ \hline
\multicolumn{1}{|l|}{$\I_{ss}$} &            &            & \checkmark &            & \checkmark  & \checkmark &            &            & \checkmark &            & \checkmark  & \checkmark \\ \hline
\multicolumn{1}{|l|}{$\I_{rb}$} &            &            &            & \checkmark &             &            &            &            &            &            &             &            \\ \hline
\multicolumn{1}{|l|}{$\I_{ro}$} &            &            &            &            &             &            &            &            & \checkmark & \checkmark &             &            \\ \hline
\multicolumn{1}{|l|}{$\I_{id}$} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark \\ \hline
\multicolumn{1}{|l|}{$\I_{ww}$} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark \\ \hline
\end{tabular}

\caption{The semantics of isolation levels on Postgres and MySQL}
\label{fig:iso-table}
\end{table}

With help of the above axiomatizations, we can now define the precise
semantics of various isolation levels on Postgres and MySQL. The table
in Table~\ref{fig:iso-table} captures the summary.
